---
title: "Probability Distribution Checks"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

Here we check the probability distribution that best fits these data. I'm following a method from http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html. As you see in that guide, there are several distributions that can be checked, for my purposes normal or lognormal are enough so I won't check the others.

## 2015 Check probability distribution fit for leaf sheath blight

```{r, results='hide', message=FALSE, warning=FALSE}

qqp(AUDPS_2015$LShB_rating, "norm")

qqp(AUDPS_2015$LShB_rating, "lnorm")

```

## 2015 Check probability distribution fit for tiller sheath blight

```{r, results='hide', message=FALSE, warning=FALSE}

AUDPS_2015$TShB_rating <- AUDPS_2015$TShB_rating + 1

qqp(AUDPS_2015$TShB_rating, "norm")

qqp(AUDPS_2015$TShB_rating, "lnorm")

```

*******

## 2016 Check probability distribution fit for leaf sheath blight

```{r, results='hide', message=FALSE, warning=FALSE}

qqp(AUDPS_2016$LShB_rating, "norm")

qqp(AUDPS_2016$LShB_rating, "lnorm")

```

## 2016 Check probability distribution fit for tiller sheath blight

```{r, results='hide', message=FALSE, warning=FALSE}

qqp(AUDPS_2016$TShB_rating, "norm")

qqp(AUDPS_2016$TShB_rating, "lnorm")

```

## Conclusions

To read the graphs, we want to select a distribution where most of the points fall between the dotted red lines. In this case, for the 2015 data, the leaf data fit a lognormal distribution best, for all others a normal distribution appears to be fine.
