---
title: "Probability Distribution Checks"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ProjectTemplate)
load.project()
```

Here we check the probability distribution that best fits these data. I'm following a method from http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html. As you see in that guide, there are several distributions that can be checked. However, in these data only normal and exponential 

## 2015

### Check probability distribution fit for tiller sheath blight incidence

#### Normal

```{r 2015_inc_norm, results='hide', message=FALSE, warning=FALSE}
AUDPS15 <- subset(AUDPS, YEAR == "2015")
f1 <- fitdist(AUDPS15$TShB_inc_AUDPS, "norm")
plot(f1)

car::qqp(AUDPS15$TShB_inc_AUDPS, "norm")
```

#### Exponential

```{r 2015_inc_exp, results='hide', message=FALSE, warning=FALSE}

f1 <- fitdist(AUDPS15$TShB_inc_AUDPS, "exp")
plot(f1)
car::qqp(AUDPS15$TShB_inc_AUDPS, "exp")
car::qqp(AUDPS15$TShB_inc_AUDPS, "lnorm")

```

*******

## 2016 

### Check probability distribution fit for tiller sheath blight incidence

#### Normal

```{r, 2016_inc_norm, results='hide', message=FALSE, warning=FALSE}
AUDPS16 <- subset(AUDPS, YEAR == "2016")
 f1 <- fitdist(AUDPS16$TShB_inc_AUDPS, "norm")
plot(f1)

car::qqp(AUDPS16$TShB_inc_AUDPS, "norm")
```

#### Exponential

```{r 2016_inc_exp, results='hide', message=FALSE, warning=FALSE}

f1 <- fitdist(AUDPS16$TShB_inc_AUDPS, "exp")
plot(f1)

car::qqp(AUDPS16$TShB_inc_AUDPS, "exp")
```




## Conclusions

To read the graphs, we want to select a distribution where most of the points fall between the dotted red lines. On the y-axis are the observations and on the x-axis the quantiles modelled by the distribution. A perfect distribution fit is represented by the solid red line and the dashed red lines are the confidence intervals of the perfect distribution fit.

### Tiller Incidence Distributions

The 2015 data fit an exponential distribution while the 2016 data fit a normal distribution.
