---
title: "Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?"
author: |
  | Adam H. Sparks^[Corresponding author: Adam.Sparks@usq.edu.au]  \textsuperscript{1\dag}, Nancy P. Castilla \textsuperscript{1}, Bjorn Ole Sander \textsuperscript{1}
  | \textsuperscript{1}International Rice Research Institute, Crop and Environmental Sciences Division, Los Baños, Laguna, Philippines, 4031
  | \textsuperscript{\dag} Current Address: University of Southern Queensland, Centre for Crop Health, West Street, Tooowoomba, Queensland, Australia, 4350
csl: "../templates/european-journal-of-plant-pathology.csl" # Insert path for the bib-style
abstract: |
  Water and nitrogen management play vital roles in rice production. However, the mismanagement of these two management practices may trigger sheath blight of rice, caused by _Rhizoctonia solani_, which is favored by wet conditions, high relative humidity, and high nitrogen fertilizer levels. To understand how different combinations of water and nitrogen management affect sheath blight epidemics, we conducted two separate split-plot experiments with a water saving (alternate wetting and drying) regime and traditional flood irrigation regime combined with differing nitrogen treatments in the dry seasons of 2015 and 2016. Disease was scored in the same way in both experiments using a sheath blight assessment scale for field evaluation developed at the International Rice Research Institute to assess the severity on infected sheaths and leaves while sheath blight incidence on tillers were counted per hill. We were unable to detect any difference in the incidence of tiller sheath blight due to irrigation, tiller and leaf sheath blight did clearly differ statistically by irrigation treatment, but leaf sheath blight severity did not. Our findings suggest that farmers can adopt water saving technologies without risking increased sheath blight incidence. We suggest that further cross-cutting research in this area is warranted. All data generated from this project and code used for analysis are available as a reproducible research compendium from <https://github.com/openplantpathology/rice_awd_pests> [@Sparks2018].
bibliography: references.bib
output:
  bookdown::word_document2:
    toc: no
    reference_docx: "../templates/template.docx"
---

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  relative_path = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  dev = c("postscript"),
  fig.width = 6.85,
  fig.height = 6.85,
  out.width = "100%",
  dpi = 300
)

library(bookdown)
library(knitr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(rice.awd.pests)
```

# Introduction {-}

Sheath blight disease (*Rhizoctonia solani* Kühn), anastomosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. Under conducive conditions the disease can cause up to 50&nbsp;\% yield losses [@Marchetti1991]. Studies by Savary et al. [-@Savary2000] estimated normal yield losses in tropical irrigated lowland rice in Asia at 5-10 &nbsp;\%. The disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009; @Srinivasachary2011] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen (N) management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

Alternate wetting and drying (AWD) is an irrigation technique for rice (*Oryza sativa* L.) developed by the International Rice Research Institute (IRRI) and its partners that saves about 15-40&nbsp;\% of irrigation water while not affecting yield [@Bouman2001; @Feng2007]. In AWD rice, fields are exposed to several dry phases during the growing season without exposing the plants to water stress. In order to avoid yield decline under AWD “safe” thresholds have been developed. Under safe AWD irrigation water is applied when the field water level reaches 15&nbsp;cm below the soil surface [@Richards2014]. Fields are furthermore kept flooded during the flowering period to avoid plant stress. As an added benefit to saving water, AWD also reduces greenhouse gas (GHG) emissions of rice fields, which are a substantial factor in the GHG budget of rice producing countries, by around 50 &nbsp;\% [@Yan2005; @Sander2015].

The AWD technology has been identified as promising climate-smart practice for different rice growing regions that can stabilize rice production in water scarce areas as well as help reduce the carbon footprint of rice production. Various countries, *e.g.*, Bangladesh, Vietnam, Thailand and Philippines, plan to widely apply AWD to rice production [@MOEF2015]. However, a change like this in the water regime in rice fields on a large scale might trigger different other effects, for example changes in disease intensity or severity.

While the effects of climate change directly on rice disease have been studied [@Duku2016 CITATION XXXX] the effects of mitigation practices have not. And because of the well documented effects of irrigation practices (CITATION XXXX) and N on sheath blight (CITATION XXXX); we established field experiments to study what effects AWD irrigation technologies and N fertilisation in combination could have on sheath blight incidence and severity in tropical rice production.

# Materials and Methods {-}

Two experiments were conducted at the International Rice Research Institute's (IRRI) Ziegler Experiment Station in Los Baños, Calabarzon, Philippines (latitude 14&deg;&nbsp;11'&nbsp;N, longitude 121&deg;&nbsp;15'&nbsp;E) in 2015 and 2016 dry seasons from December to May. In the 2016 experiment changes were made to optimize the experiment based on findings from the 2015 experiment. The differences are detailed in the following sections.

## 2015 Experiment {-}

### Experimental Design {-}

The plot design was a split-plot randomised complete block design with four replicates where irrigation was the main plot treatment and N rate was the split-plot treatment. 

The main plot treatments were alternate wetting and drying (AWD) and continuously flooded or farmers' practice as the control treatment. The main plot size was 12&nbsp;m x 12&nbsp;m (144&nbsp;m^2), with a split-plot size of 5&nbsp;m&nbsp;x&nbsp;5&nbsp;m (25&nbsp;m^2). Replication size was 12&nbsp;m&nbsp;x&nbsp;24&nbsp;m (288&nbsp;m^2) with a buffer of 1&nbsp;m per split-plot for a whole experiment size of 1,152&nbsp;m^2.

Irrigation in AWD plots was determined by the water level in plots, *i.e.*, when the water level reached 15&nbsp;cm below the soil surface irrigation water was applied to a level of 5&nbsp;cm. In flooded plots a standing water layer of 3-5&nbsp;cm was maintained throughout the growing season. 

The split-plot treatment rates of N were N0 (no N supply), N100 (100&nbsp;kg&nbsp;per&nbsp;ha applied as urea in three splits) and N120 (120&nbsp;kg&nbsp;per&nbsp;ha applied as urea in three splits) (Tab.&nbsp;\@ref(tab:N_rate)).

### Crop Establishment {-}

Trays of NSIC Rc222 were seeded on 19 December 2014. NSIC Rc222 is an inbred variety released by the Philippine Rice Research Institute (PhilRice), with a 114 day maturity when transplanted. The variety is commonly grown by farmers in the area, having good yields and moderate resistance to brown plant hopper, green leaf hopper and yellow stem borer, but susceptibility to tungro. Plots were established by manually transplanting seedlings on 9 January 2015 in hills with six to eight seedlings per hill and a distance of 20&nbsp;cm within and between rows.

### Inoculum Preparation and Application {-}

An isolate of *Rhizoctonia solani* AG1-1a previously isolated from infected rice was maintained on potato dextrose agar (PDA) medium in tubes. The isolate was transferred to 90&nbsp;mm Petri dishes containing PDA and incubated a room temperature (20 to 27&deg;C). Glass bottles of autoclaved rice grain and hull substrate were prepared and plugs of the culture were transferred from Petri dishes to the autoclaved substrate and incubated at room temperature for two weeks.

The plots were inoculated 20 days after transplanting by spreading 151&nbsp;g of inoculum over the split-plot area with a 1&nbsp;m buffer on either end (4&nbsp;m x 11&nbsp;m).

## 2016 Experiment {-}

### Experimental Design {-}

The 2016 experiment followed the same split-plot design as described for the 2015 experiment. However, the plot size was increased and due to these changes, the replicate sizes are are not equal as necessitated by the use of a larger area for the experiment. The main plot sizes were: Block 1 (B1) 21&nbsp;m x 20.5&nbsp;m (412.5 sq&nbsp;m) and Block 2 (B2) 20.25&nbsp;m x 21.6&nbsp;m (437.4&nbsp;m^2). The split-plot sizes were B1 21&nbsp;m x 10.25&nbsp;m (215.25&nbsp;m^2), B2 20.25&nbsp;m x 10.8&nbsp;m (218.7&nbsp;m^2). The replication sizes were B1 - 42&nbsp;m&nbsp;x&nbsp;20.5&nbsp;m (861&nbsp;m^2) and B2 - 40.5&nbsp;m&nbsp;x&nbsp;21.6&nbsp;m (874.8&nbsp;m^2). A buffer 0.5&nbsp;m per split-plot was used and the overall experiment size was 3471.6&nbsp;m^2.

Split-plot N rates differed from the 2015 rates with only two N treatments, N60 (60&nbsp;kg&nbsp;/&nbsp;ha as urea split into two applications) and N180 (180&nbsp;kg&nbsp;/&nbsp;ha in three splits) being applied (Tab.&nbsp;\@ref(tab:N_rate)).

### Crop Establishment {-}

Nurseries of NSIC Rc222 were established on 7 January 2016. Seedlings were manually transplanted from 20 to 22 January 2016 in hills with six to eight seedlings per hill with a distance of 20&nbsp;cm within and between rows.

### Inoculum Preparation and Application {-}

Based on the 2015 results, the inoculation methods were modified in 2016 to increase the amount of inoculum and applied to a smaller area. Plots were inoculated 41 days after transplanting using ten bottles per one sampling area (1&nbsp;m&nbsp;x&nbsp;1&nbsp;m) per plot, where one bottle contained 151&nbsp;g inoculum. A total amount of 1,510&nbsp;g of inoculum was applied to a 1&nbsp;m&nbsp;x&nbsp;1&nbsp;m area.

## Data Collection {-}

In both experiments, two sample areas, sized 1&nbsp;m&nbsp;x&nbsp;1&nbsp;m, per plot were assessed. The total number of tillers per hill and number of tillers with sheath blight (incidence) were measured for nine hills per sample area. Tiller sheath severity was measured for four tillers per hill and six leaves tiller using a disease assessment rating scale for field evaluation developed at IRRI. The scale is an unequally spaced categorical scale where 0 - no disease; 1 - trace to 1&nbsp;\% severity; 2 - 1 to 5&nbsp;\%; 3 - 5 to 15&nbsp;\%; 4 - 15 to 50&nbsp;\%; 5 - 50 to 100&nbsp;\%. Severity was considered to be the amount of leaf or sheath tissue covered by sheath blight lesions. Leaves were rated as green (living) having at least XXX&nbsp;\% living tissue or dead, defined as having less than XXX&nbsp;\% living tissue. The same rating methodology was used in both experiments. Five disease assessments were made in the 2015 experiment and four disease assessments were made in the 2016 experiment, respectively.

## Statistical Analysis {-}

Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity using the R package `agricolae` [@R2018; @deMendiburu2017]. Due to the scale used for collection, the severity data were converted to the midpoint percent value of the scale range for each value.

The correlations between tiller sheath sheath blight incidence and tiller and leaf sheath blight severity were tested using Spearman’s correlation test, `cor.test`, [@R2018]. No correlation was found, so the analysis was completed for each variable independently with no assumed interaction.

As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using Bayesian multivariate generalised linear mixed models implemented in the R package `MCMCglmm` [@MCMCglmm2010]. Six models were created where the the dependent variables were 2015 tiller sheath blight incidence, 2015 tiller sheath blight severity, 2015 leaf sheath blight severity; 2016 tiller sheath blight incidence, 2016 tiller sheath blight severity; 2016 leaf sheath blight severity, respectively. The independent variables were the irrigation management and N rate with replicate treated as a random variable. The base levels (control treatments) for the analyses for 2015 were: N rate - N0, irrigation management - flooded and for 2016 they were: N rate - N60, irrigation management - flooded.

Four MCMC chains were run for 55,000 steps with the first 5000 discarded as burn in. The remaining 50,000 samples were used to determine the posterior distribution of the parameters of the model. The `coda` [@coda2006] package was used to provide utilities to check MCMC chain convergence using visual inspection of resulting trace graphs and the Gelman-Rubin test.

# Results {-}

## 2015 Experiment

### Tiller Sheath Blight Incidence {-}

In 2015 the incidence of tiller sheath blight remained low throughout the growing season (Fig.&nbsp;\@ref(fig:fig-1)a, \@ref(fig:fig-2)a). Water management was not statistically clearly different (Fig.&nbsp;\@ref(fig:fig-3)a, \@ref(fig:fig-5)a). However, the N treatments, N100 and N120, were both statistically clearly different when compared with the control N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)a, \@ref(fig:fig-5)a). 

### Tiller Sheath Blight Severity {-}

Tiller sheath blight severity remained below 2&nbsp;\% (Fig.&nbsp;\@ref(fig:fig-2)c, \@ref(fig:fig-1)c). Both the N100 and N120 treatments were statistically clearly different than the control N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)c, \@ref(fig:fig-5)c). However, the AWD water management was not statistically clearly different from the flooded treatment (Fig.&nbsp;\@ref(fig:fig-3)c, \@ref(fig:fig-5)c).

### Leaf Sheath Blight Severity {-}

Severity of leaf sheath blight remained low, less than 0.4&nbsp;% across all treatments (Fig.&nbsp;\@ref(fig:fig-1)e, \@ref(fig:fig-2)e). Both the N100 and N120 treatments were statistically clearly different than the N0 treatment (Fig.&nbsp;\@ref(fig:fig-4)e, \@ref(fig:fig-5)e). However, the AWD water management was not statistically clearly different from the flooded treatment (Fig.&nbsp;\@ref(fig:fig-3)e, \@ref(fig:fig-5)e).

### The Interaction of Irrigation and N Management

## 2016 Experiment {-}

### Tiller Sheath Blight Incidence {-}

The changes to the inoculation methodology resulted in a higher rate of infection in 2016, with the N180 treatment reaching a maximum value of 98&nbsp;% incidence at the third observation (Fig.&nbsp;\@ref(fig:fig-2)b). The N treatment N180, was statistically clearly different than the control N60 treatment (Fig.&nbsp;\@ref(fig:fig-4)b, \@ref(fig:fig-5)b). As in the 2015 study, water management did not statistically clearly differ (Fig.&nbsp;\@ref(fig:fig-3)b, \@ref(fig:fig-5)b).

### Tiller Sheath Blight Severity {-}

As with the tiller incidence, the tiller severity increased with the changed inoculation methods with a maximum of 7.6% for the flooded treatment (Fig.&nbsp;\@ref(fig:fig-2)d). In 2016 the N180 treatment was statistically clearly different from the N60 treatment, with N180 severity being higher (Fig.&nbsp;\@ref(fig:fig-4)d, \@ref(fig:fig-5)d). The AWD water management, which was lower than the flooded treatment, was also statistically clearly different (Fig.&nbsp;\@ref(fig:fig-3)d, \@ref(fig:fig-5)d).

### Leaf Sheath Blight Severity {-}

In 2016 the neither of the treatments, N rate or water management, were statistically clearly different from the control treatments for leaf sheath blight severity (Fig.&nbsp;\@ref(fig:fig-3)f, \@ref(fig:fig-5)f, \@ref(fig:fig-4)f, \@ref(fig:fig-5)f).

### The Interaction of Irrigation and N Management

# Discussion {-}

In both experiments we were unable to detect any statistically clear effect of AWD on sheath blight that led to increases in the disease that could hinder adoption of the technology. In fact, in only one instance were we able to detect any clear effect of AWD on sheath blight. In the 2016 experiment the tiller sheath blight severity was clearly lower for the AWD treatment than flooded treatment (Fig.&nbsp;\@ref(fig:fig-3)d, \@ref(fig:fig-5)d), indicating a possible adverse effect of using AWD on tiller sheath blight severity under high sheath blight pressure.

The findings of the effects of N rates on sheath blight were as expected in both experiments. Higher rates of N caused an increase in disease incidence and severity in 2015 and in tiller incidence in 2016. However, in the 2016 experiment there were no detectable differences due to the N rate tiller and leaf severity.

The combination of the irrigation and N rates did not appear to increase the disease severity or incidence in either experiment. In 2015 there were no clear differences in the interaction of irrigation and N rate between any of the three treatments. However, in the 2016 experiment the combination of AWD-N180 was clearly different, being lower, than the flooded-N60 combination treatment.

By increasing the plot size and increasing inoculum amount applied to a smaller area, the changes made for the 2016 experiment appear to have improved the experiment. The the sheath blight incidence increased and the variability of sheath blight in the plots decreased (Fig.&nbsp;\@ref(fig:fig-1)a:b, \@ref(fig:fig-2)c:d, \@ref(fig:fig-3)a:b, \@ref(fig:fig-4)c:d).

It should be noted that the levels of leaf severity remained low (< 1 \%) throughout the growing season in all treatments for both years. As AWD can only feasibly be implemented in the dry season, the sheath blight disease levels in these studies may be lower than they would be if conducted in the wet season. However, sheath blight remains an issue in both seasons and the use of AWD would not appear to increase the disease but may be effective in reducing it in some cases.

The rice establishment method is known to affect the spread of sheath blight [@Willocquet2000]. As manual transplanting becomes less common XXXXX...

Sheath blight has been predicted to XXXXX due to the effects of climate change (XXXXXX 20XX). 

While sheath blight remains an important disease in rice production with little resistance available in current varieties, based on these studies the use of AWD technologies would not appear to be a factor that increases the incidence or severity of rice sheath blight disease.

# Notes {-}

## Acknowledgments {-}

The authors wish to thank the IRRI Biometrics group for their input in the design of the experiments and Rachel King of the University of Southern Queensland Statistical Consulting Unit for her insightful comments on the analysis of this paper. We would also like to thank Dale Amozola, Paul Escandor and P. Lapis for their technical support in the field. Lastly, we would like to remember our teammate Michael Noel, a part of the plant disease management team at IRRI and who helped design and managed the first year of the project but passed away in September 2016.

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References {-}
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

# Tables {-}

```{r N-rates}

kable(N_rates,
      caption = "Nitrogen application rates for 2015 and 2016. Five different nitrogen treatments were applied depending on the experiment. In the 2015 experiment three nitrogen rate treatments were applied: no N (N0), 100&nbsp;kgkg&nbsp;/^&nbsp;ha N (N100) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N120). In the 2016 experiment, two nitrogen rate treatments were applied: 60&nbsp;kgkg&nbsp;/^&nbsp;ha N (N60) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N180). Treatments in both years were applied in splits at: basal, tillering growth stages for all treatments and panicle initiation growth stage for treatments with three applications.")
```

##### pagebreak

# Figures {-}

##### Sheath blight progress for water management treatments

```{r fig-1, fig.cap="Sheath blight progress for 2015 and 2016 experiments. Disease progress was measured in the 2015 experiment at five assessment points and 2016 experiment at four assessment points during the dry season both years. Main plot irrigation treatments were permanently flooded (FLD) and alternate wetting and drying (AWD). Points represent the mean observations of four replications."}

p1_anot <- data.frame(
  label = c("a", "b"),
  YEAR   = c(2015, 2016),
  x = c(60, 60),
  y = c(1, 1)
)

p1 <- ggplot(data = RAW_data,
             aes(
               x = DAI,
               y = TShB_incidence
             )) +
  geom_text(
    data = p1_anot,
     mapping = aes(x = x,
                   y = y,
                   label = label)) +
  stat_summary(fun.y = "mean",
               geom = "line",
               aes(group = WMGT,
               linetype = WMGT)) +
  stat_summary(fun.y = "mean",
               geom = "point",
               aes(group = WMGT,
               shape = WMGT)) +
  scale_y_continuous(
    labels = function(x)
      paste0(x * 100),
    limits = c(0, 1),
    name = "Tiller Incidence (%)"
  ) +
  xlab("Days After Inoculation") +
  facet_grid(. ~ YEAR,
             scales = "free_x") +
  labs(linetype = "Irrigation",
       shape = "Irrigation") +
  theme_classic() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = c(x = 0.065, y = 0.76),
    legend.background = element_rect(
      color = "black",
      size = 0.5,
      linetype = "solid"
    ),
    axis.title.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold")
  )

p2_anot <- data.frame(
  label = c("c", "d"),
  YEAR   = c(2015, 2016),
  x = c(60, 60),
  y = c(10, 10)
)

p2 <- ggplot(RAW_data,
             aes(x = DAI,
                 y = PERC_TIL_ShB)) +
  geom_text(
    data = p2_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  stat_summary(fun.y = "mean",
               geom = "line",
               aes(group = WMGT,
                   linetype = WMGT)) +
  stat_summary(fun.y = "mean",
               geom = "point",
               aes(group = WMGT,
                   shape = WMGT)) +
  scale_y_continuous(name = "Tiller Severity (%)") +
  xlab("Days After Inoculation") +
  facet_grid(. ~ YEAR,
             scales = "free_x") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none",
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank())


p3_anot <- data.frame(
  label = c("e", "f"),
  YEAR   = c(2015, 2016),
  x = c(60, 60),
  y = c(10, 10)
)

p3 <- ggplot(RAW_data,
             aes(x = DAI,
                 y = PERC_LEAF_ShB)) +
  geom_text(
    data = p3_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  stat_summary(fun.y = "mean",
               geom = "line",
               aes(group = WMGT,
                   linetype = WMGT)) +
  stat_summary(fun.y = "mean",
               geom = "point",
               aes(group = WMGT,
                   shape = WMGT)) +
  scale_y_continuous(name = "Leaf Severity (%)") +
  xlab("Days After Inoculation") +
  facet_grid(. ~ YEAR,
             scales = "free_x") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank())

grid.arrange(p1, p2, p3, ncol = 1, nrow = 3)
```

##### Sheath blight progress for N treatments

```{r fig-2, fig.cap="Sheath blight incidence progress was measured in the 2015 experiment at five assessment points and 2016 experiment at four assessment points during the dry season both years. Five different N treatments were applied to split plots depending on the experiment. In the 2015 experiment three N rate treatments were applied: no N (N0), 100&nbsp;kgkg&nbsp;/^&nbsp;ha N (N100) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N120). In the 2016 experiment, two N rate treatments were applied: 60&nbsp;kgkg&nbsp;/^&nbsp;ha N (N60) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N180). Points represent the mean observations of four replications."}

p4_anot <- data.frame(
  label = c("a", "b"),
  YEAR   = c(2015, 2016),
  x = c(60, 60),
  y = c(1, 1)
)

p4 <- ggplot(data = RAW_data,
             aes(
               x = DAI,
               y = TShB_incidence
             )) +
  geom_text(
    data = p4_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  stat_summary(fun.y = "mean",
               geom = "line",
               aes(group = NRTE,
                   linetype = NRTE)) +
  stat_summary(fun.y = "mean",
               geom = "point",
               aes(group = NRTE,
                   shape = NRTE)) +
  scale_y_continuous(
    labels = function(x)
      paste0(x * 100),
    limits = c(0, 1),
    name = "Tiller Incidence (%)"
  ) +
  xlab("Days After Inoculation") +
  facet_grid(. ~ YEAR,
             scales = "free_x") +
  labs(linetype = "N Rate",
       shape = "N Rate") +
  theme_classic() +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = c(x = 0.1,
                        y = 0.7),
    legend.background = element_rect(
      color = "black",
      size = 0.5,
      linetype = "solid"
    ),
    axis.title.x = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_text(face = "bold")
  ) +
  guides(linetype = guide_legend(ncol = 2),
         shape = guide_legend(ncol = 2))


p5_anot <- data.frame(
  label = c("c", "d"),
  YEAR   = c(2015, 2016),
  x = c(60, 60),
  y = c(10, 10)
)

p5 <- ggplot(RAW_data,
             aes(x = DAI,
                 y = PERC_TIL_ShB)) +
  geom_text(
    data = p5_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  stat_summary(fun.y = "mean",
               geom = "line",
               aes(group = NRTE,
                   linetype = NRTE)) +
  stat_summary(fun.y = "mean",
               geom = "point",
               aes(group = NRTE,
                   shape = NRTE)) +
  scale_y_continuous(name = "Tiller Severity (%)") +
  xlab("Days After Inoculation") +
  facet_grid(. ~ YEAR,
             scales = "free_x") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none",
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank())


p6_anot <- data.frame(
  label = c("e", "f"),
  YEAR   = c(2015, 2016),
  x = c(60, 60),
  y = c(10, 10)
)

p6 <- ggplot(RAW_data,
             aes(x = DAI,
                 y = PERC_LEAF_ShB)) +
  geom_text(
    data = p6_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  stat_summary(fun.y = "mean",
               geom = "line",
               aes(group = NRTE,
                   linetype = NRTE)) +
  stat_summary(fun.y = "mean",
               geom = "point",
               aes(group = NRTE,
                   shape = NRTE)) +
  scale_y_continuous(name = "Leaf Severity (%)") +
  xlab("Days After Inoculation") +
  facet_grid(. ~ YEAR,
             scales = "free_x") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank())

grid.arrange(p4, p5, p6, ncol = 1, nrow = 3)
```

##### Sheath Blight Incidence Boxplot for water management

```{r fig-3, fig.cap="Sheath blight progress was measured in the 2015 at five assessment points and 2016 experiment at four points. Sheath blight incidence was rated as the number of infected tillers divided by the total number of tillers per hill and used to calculate the area under the disease progress stairs (AUDPS). Sheath blight severity was rated on an unevenly spaced categorical scale and converted to the mid-point percentage value to calculate the AUDPS. Main plot irrigation treatments were permanently flooded (FLD) and alternate wetting and drying (AWD)."}

p7_anot <- data.frame(
  label = c("a", "b"),
  YEAR   = c(2015, 2016),
  x = c(2.5, 2.5),
  y = c(15, 15)
)

p7 <- ggplot(AUDPS,
             aes(x = WMGT, y = TShB_inc_AUDPS)) +
  geom_boxplot() +
  geom_text(
    data = p7_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  ylab("Tiller Incidence AUDPS") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_text(face = "bold")) +
  facet_grid(. ~ YEAR,
             scales = "free_x")

p8_anot <- data.frame(
  label = c("c", "d"),
  YEAR   = c(2015, 2016),
  x = c(2.5, 2.5),
  y = c(32, 32)
)

p8 <- ggplot(AUDPS,
             aes(x = WMGT, y = TShB_sev_AUDPS)) +
  geom_boxplot() +
  geom_text(
    data = p8_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  ylab("Tiller Severity AUDPS") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank()) +
  facet_grid(. ~ YEAR,
             scales = "free_x")


p9_anot <- data.frame(
  label = c("e", "f"),
  YEAR   = c(2015, 2016),
  x = c(2.5, 2.5),
  y = c(10, 10)
)

p9 <- ggplot(AUDPS,
             aes(x = WMGT, y = LShB_sev_AUDPS)) +
  geom_boxplot() +
  geom_text(
    data = p9_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  xlab("Treatment") +
  ylab("Leaf Severity AUDPS") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        strip.background = element_blank(),
        strip.text.x = element_blank()) +
  facet_grid(. ~ YEAR,
             scales = "free_x")

grid.arrange(p7, p8, p9, ncol = 1, nrow = 3)
```

##### Tiller sheath blight incidence boxplot for N rate
```{r fig-4, fig.cap="Sheath blight severity progress was measured in the 2015 at five assessment points and 2016 experiment at four points. Sheath blight incidence was rated as the number of infected tillers divided by the total number of tillers per hill and used to calculate the area under the disease progress stairs (AUDPS). Sheath blight severity was rated on an unevenly spaced categorical scale and converted to the mid-point percentage value to calculate the AUDPS. Five different N treatments were applied to split plots depending on the experiment. In the 2015 experiment three N rate treatments were applied: no N (N0), 100&nbsp;kgkg&nbsp;/^&nbsp;ha N (N100) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N120). In the 2016 experiment, two N rate treatments were applied: 60&nbsp;kgkg&nbsp;/^&nbsp;ha N (N60) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N180)."}
 
p10_anot <- data.frame(
  label = c("a", "b"),
  YEAR   = c(2015, 2016),
  x = c(3.5, 2.5),
  y = c(14, 14)
)

p10 <- ggplot(AUDPS,
             aes(x = NRTE, y = TShB_inc_AUDPS)) +
  geom_boxplot() +
  geom_text(
    data = p10_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  ylab("Tiller Incidence AUDPS") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_text(face = "bold")) +
  facet_grid(. ~ YEAR,
             scales = "free_x")

p11_anot <- data.frame(
  label = c("c", "d"),
  YEAR   = c(2015, 2016),
  x = c(3.5, 2.5),
  y = c(32, 32)
)

p11 <- ggplot(AUDPS,
             aes(x = NRTE, y = TShB_sev_AUDPS)) +
  geom_boxplot() +
  geom_text(
    data = p11_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  ylab("Tiller Severity AUDPS") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_blank(),
        strip.background = element_blank(),
        strip.text.x = element_blank()) +
  facet_grid(. ~ YEAR,
             scales = "free_x")


p12_anot <- data.frame(
  label = c("e", "f"),
  YEAR   = c(2015, 2016),
  x = c(3.5, 2.5),
  y = c(7, 7)
)

p12 <- ggplot(AUDPS,
             aes(x = NRTE, y = LShB_sev_AUDPS)) +
  geom_boxplot() +
  geom_text(
    data = p12_anot,
    mapping = aes(x = x,
                  y = y,
                  label = label)) +
  xlab("Treatment") +
  ylab("Leaf Severity AUDPS") +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black"),
        strip.background = element_blank(),
        strip.text.x = element_blank()) +
  facet_grid(. ~ YEAR,
             scales = "free_x")

grid.arrange(p10, p11, p12, ncol = 1, nrow = 3)
```

##### Tiller Estimates Plots

```{r fig-5, fig.cap="Posterior means and 95&nbsp;% credible intervals for the explanatory variables used in models of alternate wetting and drying irrigations (AWD) and N rate (NRTE) on sheath blight tiller incidence and tiller severity. Five different N treatments were applied to split plots depending on the experiment. In the 2015 experiment three N rate treatments were applied: no N (N0), 100&nbsp;kgkg&nbsp;/^&nbsp;ha N (N100) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N120). In the 2016 experiment, two N rate treatments were applied: 60&nbsp;kgkg&nbsp;/^&nbsp;ha N (N60) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N180)."}

models <- readRDS("../data/derived_data/models.Rds")

par(mfcol = c(2, 2))

plot_tiller_estimates_paper(models[[1]], main_title = "2015")
mtext(text = "Tiller Incidence", side = 2, line = 6)
mtext(text = "a", side = 3, line = 0, adj = 1, cex = 0.75)

plot_tiller_estimates_paper(models[[2]])
mtext(text = "Tiller Severity", side = 2, line = 6)
mtext(text = "c", side = 3, line = 0, adj = 1, cex = 0.75)

plot_tiller_estimates_paper(models[[4]], main_title = "2016")
mtext(text = "b", side = 3, line = 0, adj = 1, cex = 0.75)

plot_tiller_estimates_paper(models[[5]])
mtext(text = "d", side = 3, line = 0, adj = 1, cex = 0.75)
```

##### Leaf estimates plots

```{r fig-6, fig.cap="Posterior means and 95&nbsp;% credible intervals for the explanatory variables used in models of alternate wetting and drying irrigations (AWD) and N rate (NRTE) and the interaction of the two treatments on sheath blight leaf severity. Five different N treatments were applied to split plots depending on the experiment. In the 2015 experiment three N rate treatments were applied: no N (N0), 100&nbsp;kgkg&nbsp;/^&nbsp;ha N (N100) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N120). In the 2016 experiment, two N rate treatments were applied: 60&nbsp;kgkg&nbsp;/^&nbsp;ha N (N60) and 180&nbsp;kgkg&nbsp;/^&nbsp;ha N (N180)."}

par(mfcol = c(1, 2))

plot_tiller_estimates_paper(models[[3]], main_title = "2015")
mtext(text = "Leaf Severity", side = 2, line = 6)
mtext(text = "a", side = 3, line = 0, adj = 1, cex = 0.75)

plot_tiller_estimates_paper(models[[6]], main_title = "2016")
mtext(text = "b", side = 3, line = 0, adj = 1, cex = 0.75)
```

##### pagebreak

### Colophon {-}

This report was generated on `r Sys.time()` using the following computational environment and dependencies: 

```{r colophon, cache = FALSE}
# which R packages and versions?
sessioninfo::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```
