---
title: "Data Preprocessing and Area Calculations"
author: "A. H. Sparks"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: TRUE
    toc_float: true
    theme: journal
vignette: >
  %\VignetteIndexEntry{Data Preprocessing and Area Calculations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(eval = FALSE)
```

# Introduction

## Data Availability

The data are openly available for download from Zenodo, [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3889800.svg)](https://doi.org/10.5281/zenodo.3889800), with full documentation (Sparks et al. 2020).
These same data are also distributed with this R package for completeness, available in the `inst/extdata` folder or just `extdata` when installed locally.

## Notes

*Note* TRT is created here as a combination of the main plot/split plot treatment.
This is used in place of plot numbers to keep the data in order.
However, it is not used in the final analysis as a treatment effect.

# Load Libraries

```{r libraries}
library("tidyverse")
library("agricolae")
library("usethis")
library("lubridate")
```

# Load Data

## 2015 Experiment

Preprocess 2015 data for all assessment dates.

```{r 2015Exp}
# Preprocess 2015 data for all assessment dates

DS2015_Raw_22DAI <-
  download.file(
    url = "https://zenodo.org/record/3889800/files/DS2015_Raw_22DAI.csv?download=1",
    mode = "wb",
    destfile = file.path(tempdir(), "DS2015_Raw_22DAI.csv")
  )

DS2015_Raw_35DAI <-
  download.file(
    url = "https://zenodo.org/record/3889800/files/DS2015_Raw_35DAI.csv?download=1",
    mode = "wb",
    destfile = file.path(tempdir(), "DS2015_Raw_35DAI.csv")
  )

DS2015_Raw_49DAI <-
  download.file(
    url = "https://zenodo.org/record/3889800/files/DS2015_Raw_49DAI.csv?download=1",
    mode = "wb",
    destfile = file.path(tempdir(), "DS2015_Raw_49DAI.csv")
  )

DS2015_Raw_62DAI <-
  download.file(
    url = "https://zenodo.org/record/3889800/files/DS2015_Raw_62DAI.csv?download=1",
    mode = "wb",
    destfile = file.path(tempdir(), "DS2015_Raw_62DAI.csv")
  )

DS2015_Raw_83DAI <-
  download.file(
    url = "https://zenodo.org/record/3889800/files/DS2015_Raw_83DAI.csv?download=1",
    mode = "wb",
    destfile = file.path(tempdir(), "DS2015_Raw_83DAI.csv")
  )

files <-
  list.files(tempdir(), pattern = "^DS2015_Raw", full.names = TRUE)

reformat <- function(files) {
  x <- read_csv(
    files,
    cols(
      col_character(),
      col_character(),
      col_character(),
      col_character(),
      col_character(),
      col_character(),
      col_double(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer(),
      col_integer()
    ),
    col_names = TRUE
  )
  x[is.na(x)] <- 0
  
  x <- arrange(x, REP, TRT)
  
  # add plot numbers
  x$PLOT <- rep(1:24, each = 18)
  
  # Calculate tiller sheath blight incidence -----------------------------------
  x <-
    x %>%
    mutate(TShB_incidence = NTShB / NTIL)
  
  TShB_incidence <-
    x %>% select(REP,
                 TRT,
                 WMGT,
                 NRTE,
                 SMPL,
                 HILL,
                 HGHT,
                 NTIL,
                 NTShB,
                 TShB_incidence)
  
  # Gather tillers--------------------------------------------------------------
  TIL <- vector(mode = "list", length = 4)
  
  for (i in 1:4) {
    y <-
      x %>%
      select(
        TRT,
        REP,
        WMGT,
        SMPL,
        NRTE,
        HILL,
        paste0("TIL", i),
        paste0("GL", i),
        paste0("DL", i),
        paste0("SHB", i),
        paste0("SLA", i),
        paste0("SLB", i),
        paste0("SLC", i),
        paste0("SLD", i),
        paste0("SLE", i),
        paste0("SLF", i)
      ) %>%
      pivot_longer(cols = starts_with("SL"),
                   names_to = "LEAF",
                   values_to = "LEAF_ShB",)
    
    names(y)[7] <- "TIL"
    names(y)[8] <- "GL"
    names(y)[9] <- "DL"
    names(y)[10] <- "TIL_ShB"
    
    y$LEAF[y$LEAF == paste0("SLA", i)] <- 1
    y$LEAF[y$LEAF == paste0("SLB", i)] <- 2
    y$LEAF[y$LEAF == paste0("SLC", i)] <- 3
    y$LEAF[y$LEAF == paste0("SLD", i)] <- 4
    y$LEAF[y$LEAF == paste0("SLE", i)] <- 5
    y$LEAF[y$LEAF == paste0("SLF", i)] <- 6
    
    TIL[[i]] <- y
  }
  
  z <- bind_rows(TIL)
  x <- left_join(TShB_incidence, z)
}

# Run reformat function for all 2015 files -------------------------------------
DS2015 <- map_df(files, reformat)

# Add dates --------------------------------------------------------------------
DS2015_22DAI <- data.frame(DATE = rep(as.Date("2015-02-12", origin = "1970-01-01"),
                                      times = nrow(DS2015) / 5),
                           # number of rows / 5 observations
                           ASMT = rep(1, times =  nrow(DS2015) / 5))

DS2015_35DAI <- data.frame(DATE =
                             rep(as.Date("2015-02-20", origin = "1970-01-01"),
                                 times = nrow(DS2015) / 5),
                           # number of rows / 5 observations
                           ASMT = rep(2, times =  nrow(DS2015) / 5))

DS2015_49DAI <- data.frame(DATE =
                             rep(as.Date("2015-03-05", origin = "1970-01-01"),
                                 times = nrow(DS2015) / 5),
                           # number of rows / 5 observations
                           ASMT = rep(3, times =  nrow(DS2015) / 5))

DS2015_62DAI <- data.frame(DATE =
                             rep(as.Date("2015-03-19", origin = "1970-01-01"),
                                 times = nrow(DS2015) / 5),
                           # number of rows / 5 observations
                           ASMT = rep(4, times =  nrow(DS2015) / 5))

DS2015_83DAI <- data.frame(DATE =
                             rep(as.Date("2015-04-01", origin = "1970-01-01"),
                                 times = nrow(DS2015) / 5),
                           # number of rows / 5 observations
                           ASMT = rep(5, times =  nrow(DS2015) / 5))

DS2015_DATES <-
  rbind(DS2015_22DAI,
        DS2015_35DAI,
        DS2015_49DAI,
        DS2015_62DAI,
        DS2015_83DAI)

DS2015 <- cbind(DS2015, DS2015_DATES)

# Add days after inoculation column
DS2015 <-
  DS2015 %>%
  mutate(DAI = case_when(ASMT == 1 ~ 14,
                         ASMT == 2 ~ 22,
                         ASMT == 3 ~ 35,
                         ASMT == 4 ~ 49,
                         ASMT == 5 ~ 62))

# Replace "N*" with a number for NRTE ------------------------------------------
DS2015$NRTE[which(DS2015$NRTE == "N0")] <- 0
DS2015$NRTE[which(DS2015$NRTE == "N1")] <- 100
DS2015$NRTE[which(DS2015$NRTE == "N2")] <- 120

# Add treatment numbers --------------------------------------------------------
DS2015$TRT <- NA
DS2015$TRT[which(DS2015$WMGT == "PDL" &
                   DS2015$NRTE == 0)] = "PDL_N0"
DS2015$TRT[which(DS2015$WMGT == "PDL" &
                   DS2015$NRTE == 100)] = "PDL_N100"
DS2015$TRT[which(DS2015$WMGT == "PDL" &
                   DS2015$NRTE == 120)] = "PDL_N120"
DS2015$TRT[which(DS2015$WMGT == "AWD" &
                   DS2015$NRTE == 0)] = "AWD_N0"
DS2015$TRT[which(DS2015$WMGT == "AWD" &
                   DS2015$NRTE == 100)] = "AWD_N100"
DS2015$TRT[which(DS2015$WMGT == "AWD" &
                   DS2015$NRTE == 120)] = "AWD_N120"

DS2015$YEAR <- year(DS2015$DATE)

DS2015$PLOT <- rep(1:24, each = 432)

# Arrange columns --------------------------------------------------------------
DS2015 <-
  DS2015 %>% select(
    YEAR,
    DATE,
    ASMT,
    DAI,
    REP,
    PLOT,
    TRT,
    SMPL,
    HILL,
    TIL,
    LEAF,
    GL,
    DL,
    TShB_incidence,
    TIL_ShB,
    LEAF_ShB,
    WMGT,
    NRTE
  ) %>%
  mutate(REP = as.integer(REP),
         HILL = as.integer(HILL)) %>% 
  ungroup()

DS2015$LEAF_ShB <- as.numeric(DS2015$LEAF_ShB)

DS2015 <- arrange(DS2015, ASMT, PLOT)

# Create new columns of days to calculate AUDPS -------------------------------

DAYS <- as.numeric(substr(files, 73, 74))

DS2015$DAYS <- c(rep(0, 10368), rep(diff(DAYS), each = 10368))

rm(files, reformat)
```

## 2016 Experiment

 Preprocess 2016 data for all assessments

1. Clean up date format and fill empty dates

2. Fill empty NTIL (for same hill, not missing values)

3. Fill empty NTShB (for same hill, not missing values)

4. Add column with number of days after inoculation (DAI)

```{r 2016exp}
# First assessment -------------------------------------------------------------
DS2016_A1 <-
  read_csv("https://zenodo.org/record/3889800/files/DS2016_Raw_1.csv?download=1")
# Fill missing values in NTIL and NTShB
DS2016_A1 <-
  DS2016_A1 %>% fill(NTIL)
DS2016_A1 <-
  DS2016_A1 %>% fill(NTShB)

# Fill dates
DS2016_A1[, 1] <- as.Date("2016-03-15", origin = "1970-01-01")
DS2016_A1$ASMT <- rep(1, nrow(DS2016_A1))
DS2016_A1$PLOT <- rep(1:16, each = 72)

# Fill days after inoculation
DS2016_A1$DAI <- rep(14, nrow(DS2016_A1))

# Second assessment -----------------------------------------------------------
DS2016_A2 <-
  read_csv("https://zenodo.org/record/3889800/files/DS2016_Raw_2.csv?download=1")
DS2016_A2 <-
  DS2016_A2 %>% fill(NTIL)
DS2016_A2 <-
  DS2016_A2 %>% fill(NTShB)

# Fill dates
DS2016_A2[, 1] <- as.Date("2016-03-29", origin = "1970-01-01")
DS2016_A2$ASMT <- rep(2, nrow(DS2016_A2))
DS2016_A2$PLOT <- rep(1:16, each = 72)

# Fill days after inoculation
DS2016_A2$DAI <- rep(28, nrow(DS2016_A2))

# Third assessment ------------------------------------------------------------
DS2016_A3 <-
  read_csv("https://zenodo.org/record/3889800/files/DS2016_Raw_3.csv?download=1")
DS2016_A3 <-
  DS2016_A3 %>% fill(NTIL)
DS2016_A3 <-
  DS2016_A3 %>% fill(NTShB)

# Fill dates
DS2016_A3[, 1] <- as.Date("2016-04-12", origin = "1970-01-01")
DS2016_A3$ASMT <- rep(3, nrow(DS2016_A3))
DS2016_A3$PLOT <- rep(1:16, each = 72)

# Fill days after inoculation
DS2016_A3$DAI <- rep(43, nrow(DS2016_A3))

# Fourth assessment ------------------------------------------------------------
DS2016_A4 <-
  read_csv("https://zenodo.org/record/3889800/files/DS2016_Raw_4.csv?download=1")
DS2016_A4 <-
  DS2016_A4 %>% fill(NTIL)
DS2016_A4 <-
  DS2016_A4 %>% fill(NTShB)

# Fill dates
DS2016_A4[, 1] <- as.Date("2016-04-26", origin = "1970-01-01")
DS2016_A4$ASMT <- rep(4, nrow(DS2016_A4))
DS2016_A4$PLOT <- rep(1:16, each = 72)

# Fill days after inoculation
DS2016_A4$DAI <- rep(59, nrow(DS2016_A4))

# Combine all assessments ------------------------------------------------------
DS2016 <- rbind(
  as.data.frame(DS2016_A1),
  as.data.frame(DS2016_A2),
  as.data.frame(DS2016_A3),
  as.data.frame(DS2016_A4)
)

# Set any NA values to 0, per Dr. Castilla on 060317 ---------------------------
# "We used to put zeroes on the data sheets, but Dr. Savary later suggested that
# we leave the cells blank because it has caused difficulties in proof reading
# the data. We also found more mistakes in data encoding when there are a lot
# of zeroes."
# Set any NA values to 0, per Dr. Castilla on 060317 ---------------------------

DS2016[is.na(DS2016)] <- 0

# Replace "N*" with a number for NRTE ------------------------------------------
DS2016$NRTE[which(DS2016$NRTE == "N1")] = 60
DS2016$NRTE[which(DS2016$NRTE == "N2")] = 180

# Add treatment numbers --------------------------------------------------------
DS2016$TRT <- NA
DS2016$TRT[which(DS2016$WMGT == "PDL" &
                   DS2016$NRTE == 60)] = "PDL_N60"
DS2016$TRT[which(DS2016$WMGT == "PDL" &
                   DS2016$NRTE == 180)] = "PDL_N180"
DS2016$TRT[which(DS2016$WMGT == "AWD" &
                   DS2016$NRTE == 60)] = "AWD_N60"
DS2016$TRT[which(DS2016$WMGT == "AWD" &
                   DS2016$NRTE == 180)] = "AWD_N180"

# gather leaf severity values in one col
DS2016 <-
  DS2016 %>%
  pivot_longer(starts_with("ShBL"),
               names_to = "LEAF",
               values_to = "LEAF_ShB")

# reclassify LEAF column to leaf numbers 1...6----------------------------------
DS2016$LEAF[DS2016$LEAF == "ShBL1"] <- 1
DS2016$LEAF[DS2016$LEAF == "ShBL2"] <- 2
DS2016$LEAF[DS2016$LEAF == "ShBL3"] <- 3
DS2016$LEAF[DS2016$LEAF == "ShBL4"] <- 4
DS2016$LEAF[DS2016$LEAF == "ShBL5"] <- 5
DS2016$LEAF[DS2016$LEAF == "ShBL6"] <- 6

# calculate tiller sheath blight percent incidence -----------------------------
DS2016 <-
  DS2016 %>%
  group_by(DATE, ASMT, DAI, REP, PLOT, TRT, WMGT, NRTE, SMPL, HILL) %>%
  mutate(TShB_incidence = NTShB / NTIL)

# rename column to match 2015 names
DS2016 <- rename(DS2016, TIL_ShB = NSHShB)

DS2016$YEAR <- year(DS2016$DATE)
DS2016$LEAF_ShB <- as.numeric(DS2016$LEAF_ShB)

DS2016 <-
  DS2016 %>% select(
    YEAR,
    DATE,
    ASMT,
    DAI,
    REP,
    PLOT,
    TRT,
    SMPL,
    HILL,
    TIL,
    LEAF,
    GL,
    DL,
    TShB_incidence,
    TIL_ShB,
    LEAF_ShB,
    WMGT,
    NRTE
  )

DS2016 <- arrange(DS2016, ASMT, PLOT)

# Create new columns of days to calculate AUDPS -------------------------------
DATE_1_2016 <- DATE_2_2016 <- DS2016$DATE
DATE_1_2016[which(DATE_1_2016 == min(DATE_1_2016))] <- NA
DATE_2_2016[which(DATE_2_2016 == max(DATE_2_2016))] <- NA

DATE_1 <- na.omit(DATE_1_2016)
DATE_2 <- na.omit(DATE_2_2016)

DAYS_2016 <- time_length(DATE_1 - DATE_2, unit = "day")
DS2016$DAYS <-
  c(rep(0, nrow(DS2016) - length(DAYS_2016)), DAYS_2016)

# clean up workspace ------------------------------------------------------------
rm(
  DS2016_A1,
  DS2016_A2,
  DS2016_A3,
  DS2016_A4,
  DATE_1,
  DATE_1_2016,
  DATE_2,
  DATE_2_2016,
  DAYS_2016
)

# Arrange columns --------------------------------------------------------------
DS2016 <-
  DS2016 %>% select(
    YEAR,
    DATE,
    ASMT,
    DAI,
    REP,
    PLOT,
    TRT,
    SMPL,
    HILL,
    TIL,
    LEAF,
    GL,
    DL,
    TShB_incidence,
    TIL_ShB,
    LEAF_ShB,
    WMGT,
    NRTE,
    DAYS
  ) %>% 
  ungroup()
```

# Combine 2015 and 2016 Data

First join the data from the two experiments to perform some formatting and reordering on the full set of data.
Then create new data frames with the reformatted data.

```{r join_experiments}
# Join the 2015 and 2016 Data into one Tibble ----------------------------------
RAW_data <- rbind(DS2015, DS2016)

# convert columns to factor ----------------------------------------------------
RAW_data <-
  RAW_data %>%
  mutate(
    YEAR = as.factor(YEAR),
    ASMT = as.factor(ASMT),
    PLOT = as.factor(PLOT),
    REP = as.factor(REP),
    TRT = as.factor(TRT),
    WMGT = as.factor(WMGT),
    NRTE = as.factor(NRTE),
    SMPL = as.factor(SMPL),
    HILL = as.factor(HILL),
    TIL = as.factor(TIL),
    LEAF = as.factor(LEAF)
  )

# reorder the treatments by level/year - low to high, 2015-2016 for graphs -----
RAW_data$NRTE <-
  fct_relevel(RAW_data$NRTE,
              "0",
              "100",
              "120",
              "60",
              "180")

RAW_data$TRT <-
  fct_relevel(
    RAW_data$TRT,
    "AWD_N0",
    "AWD_N100",
    "AWD_N120",
    "AWD_N60",
    "AWD_N180",
    "PDL_N0",
    "PDL_N100",
    "PDL_N120",
    "PDL_N60",
    "PDL_N180"
  )
```

# Calculate Area Under the Disease Progress Stairs (AUDPS)

## 2015

```{r areas2015}
DS2015 <- subset(RAW_data, YEAR == "2015")

# 2015 Tiller Incidence setup --------------------------------------------------
TShB_incidence_15 <-
  DS2015 %>%
  group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  summarise(PLOT_TShB_incidence = mean(TShB_incidence, na.rm = TRUE),
            .groups = "keep") %>%
  arrange(PLOT)

TShB_incidence_wide <-
  reshape2::dcast(TShB_incidence_15, PLOT ~ ASMT, value.var = "PLOT_TShB_incidence")

# 2015 Tiller Severity setup ---------------------------------------------------
TShB_percent_15 <-
  DS2015 %>%
  group_by(YEAR, REP, TRT, PLOT, ASMT, DAI, DAYS) %>%
  summarise(PLOT_TShB_percent = mean(TIL_ShB, na.rm = TRUE),
            .groups = "keep") %>%
  arrange(PLOT)

TShB_percent_wide <-
  reshape2::dcast(TShB_percent_15, PLOT ~ ASMT, value.var = "PLOT_TShB_percent")

# 2015 Leaf Severity setup -----------------------------------------------------
LShB_percent_15 <-
  DS2015 %>%
  group_by(YEAR, REP, TRT, PLOT, ASMT, DAI, DAYS) %>%
  summarise(PLOT_LShB_percent = mean(LEAF_ShB, na.rm = TRUE),
            .groups = "keep") %>%
  arrange(PLOT)

LShB_percent_wide <-
  reshape2::dcast(LShB_percent_15, PLOT ~ ASMT, value.var = "PLOT_LShB_percent")

# 2015 AUDPS data frames -------------------------------------------------------
TShB_incidence_15_AUDPS <-
  audps(evaluation = TShB_incidence_wide[, 2:6],
        dates = pull(TShB_incidence_15[1:5, 6]))

TShB_percent_15_AUDPS <-
  audps(evaluation = TShB_percent_wide[, 2:6],
        dates = pull(TShB_percent_15[1:5, 6]))

LShB_percent_15_AUDPS <-
  audps(evaluation = LShB_percent_wide[, 2:6],
        dates = pull(LShB_percent_15[1:5, 6]))

AUDPS_15 <-
  as_tibble(
    cbind(
      PLOT = 1:24,
      TShB_incidence_15_AUDPS,
      TShB_percent_15_AUDPS,
      LShB_percent_15_AUDPS
    )
  )

# Add YEAR/REP/TRT/ASMT cols to AUDPS values via TShB_incidence_15 df ----------------
AUDPS_15$PLOT <- as.character(AUDPS_15$PLOT)
TShB_incidence_15$PLOT <- as.character(TShB_incidence_15$PLOT)

ShB_15 <- left_join(TShB_incidence_15, AUDPS_15, by = "PLOT") %>%
  rename(
    TShB_incidence_AUDPS = TShB_incidence_15_AUDPS,
    TShB_percent_AUDPS = TShB_percent_15_AUDPS,
    LShB_percent_AUDPS = LShB_percent_15_AUDPS
  ) %>% 
  ungroup()
```

## 2016

```{r area_2016, message=FALSE}
DS2016 <- subset(RAW_data, YEAR == "2016")

# 2016 Tiller Incidence setup --------------------------------------------------
TShB_incidence_16 <-
  DS2016 %>%
  group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  summarise(PLOT_TShB_incidence = mean(TShB_incidence, na.rm = TRUE),
            .groups = "keep") %>%
  arrange(PLOT)

TShB_incidence_wide <-
  reshape2::dcast(TShB_incidence_16, PLOT ~ ASMT, value.var = "PLOT_TShB_incidence")

# 2016 Tiller Severity setup ---------------------------------------------------
TShB_percent_16 <-
  DS2016 %>%
  group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  summarise(PLOT_TShB_percent = mean(TIL_ShB, na.rm = TRUE),
            .groups = "keep") %>%
  arrange(PLOT)

TShB_percent_wide <-
  reshape2::dcast(TShB_percent_16, PLOT ~ ASMT, value.var = "PLOT_TShB_percent")

# 2016 Leaf Severity setup -----------------------------------------------------
LShB_percent_16 <-
  DS2016 %>%
  group_by(YEAR, REP, TRT, PLOT, ASMT, DAYS) %>%
  summarise(PLOT_LShB_percent = mean(LEAF_ShB, na.rm = TRUE),
            .groups = "keep") %>%
  arrange(PLOT)

LShB_percent_wide <-
  reshape2::dcast(LShB_percent_16, PLOT ~ ASMT, value.var = "PLOT_LShB_percent")

# 2016 AUDPS -------------------------------------------------------------------
TShB_incidence_16_AUDPS <-
  audps(evaluation = TShB_incidence_wide[, 2:5],
        dates = pull(TShB_incidence_16[1:4, 6]))

TShB_percent_16_AUDPS <-
  audps(evaluation = TShB_percent_wide[, 2:5],
        dates = pull(TShB_percent_16[1:4, 6]))

LShB_percent_16_AUDPS <-
  audps(evaluation = LShB_percent_wide[, 2:5],
        dates = pull(LShB_percent_16[1:4, 6]))


AUDPS_16 <-
  as_tibble(
    cbind(
      PLOT = 1:16,
      TShB_incidence_16_AUDPS,
      TShB_percent_16_AUDPS,
      LShB_percent_16_AUDPS
    )
  )

# Add YEAR/REP/TRT/ASMT cols to AUDPS values via TShB_incidence_16 df ----------------
AUDPS_16$PLOT <- as.character(AUDPS_16$PLOT)
TShB_incidence_16$PLOT <- as.character(TShB_incidence_16$PLOT)

ShB_16 <- left_join(TShB_incidence_16, AUDPS_16, by = "PLOT") %>%
  ungroup() %>%
  rename(
    TShB_incidence_AUDPS = TShB_incidence_16_AUDPS,
    TShB_percent_AUDPS = TShB_percent_16_AUDPS,
    LShB_percent_AUDPS = LShB_percent_16_AUDPS
  ) %>% 
  ungroup()
```

# Merge AUDPS Year Data

```{r merge_areas}
AUDPS <- bind_rows(ShB_15, ShB_16) %>%
  select(-PLOT_TShB_incidence)

AUDPS <- separate(
  data = AUDPS,
  col = TRT,
  sep = "_",
  into = c("WMGT", "NRTE")
)

AUDPS <- mutate_at(.tbl = AUDPS,
                   .funs = factor,
                   .vars = c("WMGT", "NRTE"))

AUDPS$NRTE <- fct_relevel(AUDPS$NRTE,
                          "N0",
                          "N100",
                          "N120",
                          "N60",
                          "N180")

AUDPS$WMGT <-
  fct_relevel(AUDPS$WMGT,
              "PDL",
              "AWD")
```

# N rates

Create a `tibble` of N rates and time of application used in this study.

```{r N_rates, message=FALSE}
Year <- as.integer(c(2015,
                     2015,
                     2015,
                     2016,
                     2016))

`Total N (kg/ha)` <- as.integer(c(0,
                                  100,
                                  120,
                                  60,
                                  180))

`Basal N (kg/ha)` <- as.integer(c(0,
                                  60,
                                  60,
                                  30,
                                  60))

`Tillering N (kg/ha)` <- as.integer(c(0,
                                      20,
                                      30,
                                      30,
                                      60))

`Panicle Initiation N (kg/ha)` <- as.integer(c(0,
                                               20,
                                               30,
                                               0,
                                               60))

N_rates <- cbind(
  Year,
  `Total N (kg/ha)`,
  `Basal N (kg/ha)`,
  `Tillering N (kg/ha)`,
  `Panicle Initiation N (kg/ha)`
)

row.names(N_rates) <- c("N0", "N100", "N120", "N60", "N180")
```

# Save Data

Save the data in this R package for analysis in the next vignette articles.

```{r save}
if (!dir.exists("data")) {
  dir.create("data", recursive = TRUE)
}

use_data(RAW_data,
         compress = "bzip2",
         overwrite = TRUE)

use_data(AUDPS,
         compress = "bzip2",
         overwrite = TRUE)

use_data(N_rates,
         compress = "bzip2",
         overwrite = TRUE
)
```

# References

Sparks, Adam H, Castilla, Nancy P, & Sander, Bjoern Ole. (2020). Data for "Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?" [Data set]. Zenodo. [DOI: 10.5281/zenodo.3889800](http://doi.org/10.5281/zenodo.3889800)
