---
title: "Visualise Data"
author: "A. H. Sparks"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    theme: journal
vignette: >
  %\VignetteIndexEntry{Visualise_Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Libraries

Load packages necessary to conduct this work.

```{r libraries, message=FALSE, warning=FALSE, results="hide"}
library("rice.awd.shb")
library("dplyr")
library("ggplot2")
theme_set(theme_minimal())
```

# Histogram plot of leaf sheath blight severity data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = LEAF_ShB)) +
  geom_histogram(aes(fill = YEAR), stat = "count") +
  facet_grid(WMGT ~ NRTE) +
  ggtitle("Leaf Sheath Blight Severity Rating")
```

# Histogram plot of tiller sheath blight severity data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = TIL_ShB)) +
  geom_histogram(aes(fill = YEAR), stat = "count") +
  facet_grid(WMGT ~ NRTE) +
  ggtitle("Tiller Sheath Blight Severity Rating")
```

# Density plot of tiller sheath blight incidence data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = TShB_incidence)) +
  geom_density(aes(fill = YEAR, colour = YEAR)) +
  geom_rug() +
  scale_linetype(name = "Year") +
  facet_grid(WMGT ~ NRTE) +
  ggtitle("Tiller Sheath Blight Incidence") +
theme(axis.text.x = element_text(size = 8,
                                 angle = 45,
                                 hjust = 1))
```

# Histogram plot of green leaf data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = GL, linetype = YEAR)) +
  geom_histogram(aes(fill = YEAR), position = "dodge") +
  xlab("Count") +
  ggtitle("Green Leaves per Tiller")
```

# Histogram plot plot of dead leaf data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = DL, linetype = YEAR)) +
  geom_histogram(aes(fill = YEAR), position = "dodge") +
  xlab("Count") +
  ggtitle("Dry Leaves per Tiller")
```

# Dot plot of leaf sheath blight severity data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = ASMT, y = LEAF_ShB)) +
  geom_point(position = position_jitter(width = 0.05),
             aes(colour = YEAR), size = 0.75) +
  xlab("Assessment") +
  ylab("Rating") +
  ggtitle("Leaf Sheath Blight Severity at Each Assessment") +
  facet_grid(. ~ YEAR) +
  theme(strip.text.x = element_text(angle = 90))
```

# Dot plot of tiller sheath blight severity data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = ASMT, y = TIL_ShB)) +
  geom_point(position = position_jitter(width = 0.05),
             aes(colour = YEAR), size = 0.75) +
  xlab("Assessment") +
  ylab("Rating") +
  facet_grid(. ~ YEAR) +
  ggtitle("Tiller Sheath Blight Severity at Each Assessment") +
  theme(strip.text.x = element_text(angle = 90))
```

# Dot plot of tiller sheath blight incidence data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = ASMT, y = TShB_incidence,
                     group = TShB_incidence)) +
  geom_point(position = position_jitter(width = 0.05),
             aes(colour = YEAR), size = 0.75) +
  xlab("Assessment") +
  ylab("Incidence") +
  facet_grid(. ~ YEAR) +
  ggtitle("Tiller Sheath Blight Severity Rating Over Time") +
  theme(strip.text.x = element_text(angle = 90)) +
  xlab("Assessment") +
  ylab("Incidence") +
  ggtitle("Tiller Sheath Blight at Each Assessment")
```

# Violin/dot plots of tiller sheath blight incidence data

```{r, cache=FALSE}
ggplot(RAW_data, aes(x = WMGT, y = TShB_incidence)) +
  geom_point(position = position_jitter(width = 0.2),
             aes(colour = YEAR), size = 0.1) +
  geom_violin(fill = NA) +
  theme(axis.text.x = element_text(size = 8,
                                   angle = 45,
                                   hjust = 1)) +
  xlab("Treatment") +
  ylab("Incidence per Hill") +
  facet_grid(. ~ NRTE) +
  ggtitle("Tiller Sheath Blight Incidence")
```

# Box/dotplots of tiller sheath blight incidence AUDPS data

```{r, cache=FALSE}

ggplot(AUDPC_S, aes(x = NRTE, y = TShB_inc_AUDPS)) +
  geom_boxplot(aes(fill = YEAR)) +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  ggtitle("Tiller Sheath Blight Incidence")
```

```{r, cache=FALSE}
ggplot(AUDPC_S, aes(x = WMGT, y = TShB_inc_AUDPS)) +
  geom_boxplot(aes(fill = YEAR)) +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  facet_grid(. ~ YEAR) +
  ggtitle("Tiller Sheath Blight Incidence")
```


# Box/dotplots of leaf sheath blight severity AUDPS

```{r, cache=FALSE}
ggplot(AUDPC_S, aes(x = NRTE, y = LShB_percent_AUDPS)) +
  geom_boxplot(aes(fill = YEAR)) +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  ggtitle("Leaf Sheath Blight Severity")
```

```{r, cache=FALSE}
ggplot(AUDPC_S, aes(WMGT, y = LShB_percent_AUDPS)) +
  geom_boxplot(aes(fill = YEAR)) +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  facet_grid(. ~ YEAR) +
  ggtitle("Leaf Sheath Blight Severity")
```

# Box/dotplots of tiller sheath blight severity AUDPS data

```{r, cache=FALSE}
ggplot(AUDPC_S, aes(x = WMGT, y = TShB_percent_AUDPS)) +
  geom_boxplot(aes(fill = WMGT)) +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  facet_grid(. ~ YEAR) +
  ggtitle("Tiller Sheath Blight Severity")
```

```{r, cache=FALSE}
ggplot(AUDPC_S, aes(x = NRTE, y = TShB_percent_AUDPS)) +
  geom_boxplot(aes(fill = YEAR)) +
  geom_point() +
  xlab("Treatment") +
  ylab("AUDPS") +
  ggtitle("Tiller Sheath Blight Severity")
```

# Line graph of leaf sheath blight severity data

```{r, cache=FALSE}
RAW_data %>% group_by(YEAR, WMGT, NRTE, ASMT) %>%
  summarise(value = mean(PERC_LEAF_ShB)) %>%
  ggplot(aes(x = ASMT,
             y = value,
             color = NRTE,
             group = NRTE)) +
  geom_line() +
  xlab("Assessment") +
  ylab("Severity (%)") +
  facet_grid(WMGT ~ YEAR) +
  ggtitle("Leaf Sheath Blight")
```

# Line graph of tiller sheath blight severity data

```{r, cache=FALSE}
RAW_data %>% group_by(YEAR, WMGT, NRTE, ASMT) %>%
  summarise(value = mean(PERC_TIL_ShB)) %>%
  ggplot(aes(x = ASMT,
             y = value,
             color = NRTE,
             group = NRTE)) +
  geom_line() +
  xlab("Assessment") +
  ylab("Severity (%)") +
  facet_grid(WMGT ~ YEAR) +
  ggtitle("Tiller Sheath Blight")
```

# Line graph of tiller sheath blight incidence data

```{r, cache=FALSE}
RAW_data %>% group_by(YEAR, WMGT, NRTE, ASMT) %>%
  summarise(value = mean(TShB_incidence)) %>%
  ggplot(aes(x = ASMT,
             y = value,
             color = NRTE,
             group = NRTE)) +
  geom_line() +
  xlab("Assessment") +
  ylab("Incidence (%)") +
  facet_grid(WMGT ~ YEAR) +
  ggtitle("Tiller Sheath Blight")
```
