---
title: "2015 Tiller Sheath Blight Incidence Analysis"
author: "A. H. Sparks"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: journal
vignette: >
  %\VignetteIndexEntry{2015 Tiller Sheath Blight Incidence Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

For this analysis, I've elected to use [`rstanarm`](https://cran.r-project.org/package=rstanarm) for a Bayesian analysis.
When comparing the treatments in the analysis, the model compares each treatment to the base (control) treatment.

The base levels for this analysis are:

- `NRTE:0`

- `WMGT:PDL`

## Setup

Load libraries and detect cores.

```{r setup, message=FALSE}
library("rice.awd.shb")
library("rstanarm")
library("bayestestR")
library("bayesplot")
library("ggplot2")
library("pander")

options(mc.cores = parallel::detectCores())
```

The `AUDPS` object is loaded with the `rice.awd.shb` R package. To see how the AUDPS data were generated from the original raw data, see the [vignette detailing the data processing](a02_Data_preprocessing.html) file.
However, because it is a `tibble()` and the treatments exist in a single column for graphing the raw data, this object needs a few minor changes to be usable for the analysis.

Create individual data frames for the analysis.

```{r create_analysis_df, cache=FALSE, message=FALSE, warning=FALSE}
# create 2015 data frame
AUDPS_2015 <- as.data.frame(AUDPS[AUDPS$YEAR == 2015,])
AUDPS_2015 <- droplevels(AUDPS_2015)

# relevel factors for easier interpretation of analysis
AUDPS_2015 <- within(AUDPS_2015, NRTE <- relevel(NRTE, ref = "N0"))
AUDPS_2015 <- within(AUDPS_2015, WMGT <- relevel(WMGT, ref = "PDL"))

# add main plot effect
AUDPS_2015$main_plot <- AUDPS_2015$WMGT
```

Now that the `AUDPS_2015` and `AUDPS_2016` data frames exist, we can start the analysis.

## Tiller Sheath Blight Incidence Model

### Model Structure

Tiller sheath blight incidence AUDPS, `TShB_incidence_AUDPS` is the response variable.
Water management, `WMGT`, and nitrogen rate, `NRTE`, are fixed effects.
Replicate, `REP`, is treated as a random effect with the main plot effect, `WMGT` nested within it.

```{r 2015_TShB_incidence, cache=FALSE, message=TRUE, warning=TRUE, results="hide"}
m1_prior <- normal(location = 0,
                   scale = 10)

m1 <-
  stan_lmer(
    formula = TShB_incidence_AUDPS ~ WMGT + NRTE + (1 | REP/main_plot),
    data = AUDPS_2015,
    seed = 27,
    adapt_delta = 0.99,
    prior = m1_prior
  )
```

### Diagnostics

```{r m1-diagnostics, cache=FALSE, message=FALSE, warning=FALSE}
mcmc_trace(m1)

pp_check(m1, plotfun = "stat_2d", stat = c("mean", "sd"))
```

### Summary

```{r m1-summary, message=FALSE, warning=FALSE}
summary(m1)

et <- equivalence_test(m1)

plot(et) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()

pander(et)
```

## Save Model Information

Lastly, save the model information to be used in discussing the results.

```{r save_model}
saveRDS(m1, "../analysis/data/derived_data/TInc15.Rds",
        compress = "bzip2")
```
