---
title: "2016 Tiller Sheath Blight Incidence Analysis"
author: "A. H. Sparks"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: journal
vignette: >
  %\VignetteIndexEntry{2015 Tiller Sheath Blight Incidence Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


For this analysis, I've elected to use [`MCMCglmm`](https://cran.r-project.org/package=MCMCglmm) for a Bayesian analysis.
When comparing the treatments in the analysis, the model compares each treatment to the base (control) treatment.


The base levels for this analysis are:

- `NRTE:60`

- `WMGT:PDL`

## Setup

Load libraries and set seed for reproducibility.

```{r set_seed, message=FALSE}
library("rice.awd.shb")
library("MCMCglmm")
library("bayestestR")
library("ggplot2")

set.seed(27)
```

The `AUDPS` object is loaded with the `rice.awd.shb` R package. To see how the AUDPS data were generated from the original raw data, see the [vignette detailing the data processing](a02_Data_preprocessing.html) file.
However, because it is a `tibble()` and the treatments exist in a single column for graphing the raw data, this object needs a few minor changes to be usable for the analysis.

Create individual data frames for the analysis.

```{r create_analysis_df, cache=FALSE, message=FALSE, warning=FALSE}
# create 2016 data frame
AUDPS_2016 <- as.data.frame(AUDPS[AUDPS$YEAR == 2016,])
AUDPS_2016 <- droplevels(AUDPS_2016)

# relevel factors for easier interpretation of analysis
AUDPS_2016 <- within(AUDPS_2016, NRTE <- relevel(NRTE, ref = "N60"))
AUDPS_2016 <- within(AUDPS_2016, WMGT <- relevel(WMGT, ref = "PDL"))
```

## Tiller Sheath Blight Incidence Model

### Model Structure

Tiller sheath blight incidence AUDPS, `TShB_incidence_AUDPS` is the response variable.
Water management, `WMGT`, and nitrogen rate, `NRTE`, are fixed effects.
Replicate, `REP`, is treated as a random effect.

```{r 2016_TShB_severity, cache=FALSE, message=TRUE, warning=TRUE, results="hide"}
m4_prior <- list(R = list(V = 1, nu = 0.02),
                 G = list(G1 = list(
                   V = 1, nu = 0.02, alpha.V = 1000
                 )))

m4 <- MCMCglmm(
  TShB_incidence_AUDPS ~ WMGT * NRTE,
  ~ REP,
  data    = AUDPS_2016,
  prior = m4_prior,
  nitt = 60000,
  thin = 10
)
```

### Diagnostics and Summary

```{r m5-diagnostics, cache=FALSE, message=FALSE, warning=FALSE, fig.height = 9, fig.width = 6}
plot(m4)

summary(m4)

pd <- p_direction(m4)

pander(pd)

plot(pd) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```

# Save Model Information

Lastly, save the model information to be used in discussing the results.

```{r save_model}
saveRDS(m4, "../analysis/data/derived_data/TInc16.Rds",
        compress = "bzip2")
```
